---
title: "hw 2 242"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
setwd("C:/Users/Cynthiali0621/Desktop")
```

```{r, warning=FALSE,  message=FALSE}
library(dplyr)
library(tidyverse)
framingham_train <- read_csv("framingham_train_sp24.csv")
framingham_test <- read_csv("framingham_test_sp24.csv")
```


a) i)

```{r}
model <- glm(TenYearCHD ~ ., family = binomial, data = framingham_train)

summary(model)
```




$$
P(Y = 1 \mid X) = \frac{1}{1 + e^{-(\beta_0 + \beta_1X_1 + \beta_2X_2 + \dots + \beta_kX_k)}}
$$

Where:
- \( Y = 1 \) indicates that the patient experiences coronary heart disease within 10 years of their first examination

- \( X_1, X_2, \dots, X_k \) are the features(predictors)

- \( \beta_0, \beta_1, \dots, \beta_k \) are the coefficients from the logistic regression model.


P(TenYearCHD = 1 | X) = 1 / (1 + exp( -( -9.082582 
+ 0.421662 * male 
+ 0.071141 * age 
- 0.121964 * educationHighSchoolGED 
- 0.087004 * educationSomeCollege 
+ 0.051935 * educationSomeHighSchool 
+ 0.150589 * currentSmoker 
+ 0.024958 * cigsPerDay 
+ 0.287261 * BPMeds 
+ 0.797512 * prevalentStroke 
+ 0.221532 * prevalentHyp 
- 0.048145 * diabetes 
+ 0.002785 * totChol 
+ 0.009158 * sysBP 
+ 0.007651 * diaBP 
+ 0.005821 * BMI 
- 0.004216 * heartRate 
+ 0.008203 * glucose) ))


\newpage

a) ii)

I picked age since it has a very low p-value(and positive), where indicates it is a really useful predictors.


coefficient of age is  0.071141


odd ratio for age  = e^(0.071141) = 1.07373261149 = 1.0737


If age increase by 1 unit(assuming all remaining predictors remain unchanged), log(odds) changed by 0.071141. In other words, if age increase by 1 unit, the odds of developing CHD increase by approximately 7.37%.


\newpage

a) iii)

```{r}
library(magick)
img <- image_read("a_iii.jpg")

# Display the image
print(img)
```



\newpage

a) iv)
```{r}
thres <- 0.1105
#predict probabilities for the test set
test_predictions <- predict(model, newdata = framingham_test, type = "response")

#classify based on the threshold
test_predicted_class <- ifelse(test_predictions >= thres, 1, 0)

#confusion matrix
conf_matrix <- table(test_predicted_class, framingham_test$TenYearCHD)

#Accuracy
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)

#True Positive Rate
TPR <- conf_matrix[2,2] / (conf_matrix[2,2] + conf_matrix[1,2])

#False Positive Rate
FPR <- conf_matrix[2,1] / (conf_matrix[2,1] + conf_matrix[1,1])

list(accuracy = accuracy, TPR = TPR, FPR = FPR)

```


We classified each patient as either high risk (CHD) or low risk based on the threshold of 0.1105. For example, if a patient's predicted probability of developing coronary heart disease (CHD) within 10 years of their first examination is greater than or equal to 11.05%, it is more effective to prescribe the medication. If the probability is less than 11.05%, the medication is not prescribed.

Accuracy indicates the proportion of correct predictions in classifying patients as either high risk (CHD) or low risk based on the threshold. An accuracy of 0.5692 means that the model correctly predicted the risk category for approximately 57% of the patients in the test set. 

True Positive Rate is the proportion of actual positive cases (people who actually experience CHD within 10 years) that are correctly predicted by the model as high risk.A TPR of 77.5% means that 77.5% of the people who actually experienced CHD (the true positives) were correctly identified as being at high risk by the model.

False Positive Rate: how often the model incorrectly identifies low-risk patients as high-risk, which means  46.59% of the people who will not experience CHD within 10 years were incorrectly classified as high risk by the model.

\newpage

a) v)

```{r}
cost_without_medication <- function(p) {
  return(p * 955000)
}

cost_with_medication <- function(p) {
  return(0.1 * p * 1050000 + (1 - 0.1 * p) * 95000)
}

#expected cost without adjusting for treatment
framingham_test$expected_cost_no_treatment <- ifelse(test_predictions >= 0.1105, 
                                                     cost_without_medication(test_predictions), 
                                                     0)

#expected cost with treatment adjustment
framingham_test$expected_cost_with_treatment <- ifelse(test_predictions >= 0.1105, 
                                                       cost_with_medication(test_predictions), 
                                                       cost_without_medication(test_predictions))

#average expected cost per patient without adjusting for treatment
average_cost_no_treatment <- mean(framingham_test$expected_cost_no_treatment)

#average expected cost per patient with treatment adjustment
average_cost_with_treatment <- mean(framingham_test$expected_cost_with_treatment)

list(average_cost_no_treatment = average_cost_no_treatment, 
     average_cost_with_treatment = average_cost_with_treatment)

```


Yes, this assumption is reasonable because the data was collected before the option of prescribing the medication was available. Without adjusting for treatment, the expected cost per patient is 115097.50. With treatment adjustment, the expected cost per patient drops to 87841.34.


\newpage


a) vi)

```{r}

#for this question, I have asked chatgpt on how to built the matix, because my own matrix always occurs an error(OpenAI, 2024)

library(readr)

baseline_predictions <- rep(0, nrow(framingham_test))

conf_matrix <- table(Predicted = baseline_predictions, Actual = framingham_test$TenYearCHD)

TN <- ifelse(dim(conf_matrix)[1] >= 1 && dim(conf_matrix)[2] >= 1, conf_matrix[1, 1], 0)  
FN <- ifelse(dim(conf_matrix)[1] >= 1 && dim(conf_matrix)[2] >= 2, conf_matrix[1, 2], 0)  
FP <- 0 
TP <- 0  

#Accuracy
accuracy_baseline <- (TP + TN) / sum(conf_matrix)

#True Positive Rate
TPR_baseline <- ifelse((TP + FN) > 0, TP / (TP + FN), 0)

#False Positive Rate
FPR_baseline <- ifelse((FP + TN) > 0, FP / (FP + TN), 0)

#expected cost based on actual CHD outcomes
#If TenYearCHD = 1, the cost is $955,000; if TenYearCHD = 0, the cost is $0
framingham_test$expected_cost_baseline <- framingham_test$TenYearCHD * 955000

#average expected cost per patient for the baseline model
average_cost_baseline <- mean(framingham_test$expected_cost_baseline)

list(
  accuracy_baseline = accuracy_baseline,
  TPR_baseline = TPR_baseline,
  FPR_baseline = FPR_baseline,
  average_cost_baseline = average_cost_baseline
)


```


\newpage


a) vii)


```{r}
intercept <- -9.082582
beta_male <- 0.421662
beta_age <- 0.071141
beta_educationHighSchoolGED <- -0.121964
beta_educationSomeCollege <- -0.087004
beta_educationSomeHighSchool <- 0.051935
beta_currentSmoker <- 0.150589
beta_cigsPerDay <- 0.024958
beta_BPMeds <- 0.287261
beta_prevalentStroke <- 0.797512
beta_prevalentHyp <- 0.221532
beta_diabetes <- -0.048145
beta_totChol <- 0.002785
beta_sysBP <- 0.009158
beta_diaBP <- 0.007651
beta_BMI <- 0.005821
beta_heartRate <- -0.004216
beta_glucose <- 0.008203

male <- 0                     
age <- 39
educationHighSchoolGED <- 1    
educationSomeCollege <- 0     
educationSomeHighSchool <- 0   
currentSmoker <- 1             
cigsPerDay <- 6
BPMeds <- 0                    
prevalentStroke <- 0           
prevalentHyp <- 0              
diabetes <- 1                 
totChol <- 230
sysBP <- 110
diaBP <- 50
BMI <- 28
heartRate <- 72
glucose <- 80

linear_combination <- intercept + 
  beta_male * male + 
  beta_age * age + 
  beta_educationHighSchoolGED * educationHighSchoolGED + 
  beta_educationSomeCollege * educationSomeCollege + 
  beta_educationSomeHighSchool * educationSomeHighSchool + 
  beta_currentSmoker * currentSmoker + 
  beta_cigsPerDay * cigsPerDay + 
  beta_BPMeds * BPMeds + 
  beta_prevalentStroke * prevalentStroke + 
  beta_prevalentHyp * prevalentHyp + 
  beta_diabetes * diabetes + 
  beta_totChol * totChol + 
  beta_sysBP * sysBP + 
  beta_diaBP * diaBP + 
  beta_BMI * BMI + 
  beta_heartRate * heartRate + 
  beta_glucose * glucose

predicted_prob <- 1 / (1 + exp(-linear_combination))

predicted_prob

```

2.58% is less than the threshold of 11.05%, which means , the patient is considered low risk for experiencing coronary heart disease within 10 years of their first examination, so the physician should not prescribe the preventive medication

\newpage

b)

```{r}
library(pROC)

test_predictions <- predict(model, newdata = framingham_test, type = "response")

actual_outcomes_test <- framingham_test$TenYearCHD

roc_curve <- roc(actual_outcomes_test, test_predictions)

plot(roc_curve, col = "blue", main = "ROC Curve for Logistic Regression Model")

auc_value <- auc(roc_curve)
auc_value

```

Area under the curve: 0.71 indicates the model is performing relatively good in predicting CHD risk.AUC and ROC shows the model has reasonably high true positive rate and low false positive rates

Decision-makers can use the ROC curve to choose and study different thresholds to determine when it is most beneficial to prescribe medication, because the curve shows the trade-off between  TPR and FPR. For example, they could choose a threshold that maximizes TPR and min those midication that is uncessary for people who do not need it. 

\newpage

c)

```{r}
img1 <- image_read("c.jpg")

# Display the image
print(img1)
```

\newpage

d)

1. The analysis assumes an amount of co-payment. In reality, low-income patients may not have ability to pay the co-payment. My suggestion would be taht the insurance company could adjust the co-payment based on the patient’s ability to pay rather than a fixed amount. 

2. The threshold and all assumptions are based on cost consideration, but in reality, we should ensure that patients should fully understand their risk such as side-effects. We should provide all details for patients about the disease and all effect of treatment rather than just the cost concerns. 

