---
title: "hw4 242"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(caret)
library(pROC)
library(rpart)
library(tidyverse)
library(randomForest)
library(MASS)
library(gbm)
```

```{r}
setwd("C:/Users/Cynthiali0621/Desktop")
train_data <- read_csv("Letters_train_2024.csv") 
test_data <- read_csv("Letters_test_2024.csv") 


```

```{r}
train_data <- train_data %>% mutate(isB = ifelse(letter == "B", "Yes", "No"))
test_data <- test_data %>% mutate(isB = ifelse(letter == "B", "Yes", "No"))
```

1.(a)
```{r}
baseline_accuracy <- mean(test_data$isB == "No")
cat("1 (a) Baseline accuracy:", baseline_accuracy, "\n")
```

1.(b)

```{r}
train_data$isB_numeric <- ifelse(train_data$isB == "Yes", 1, 0)
test_data$isB_numeric <- ifelse(test_data$isB == "Yes", 1, 0)
logistic_model <- glm(isB_numeric ~ . -letter -isB, data = train_data, family = "binomial")

#Predict on test data
logistic_preds <- predict(logistic_model, newdata = test_data, type = "response")
logistic_preds_binary <- ifelse(logistic_preds > 0.5, 1, 0)

#Logistic regression accuracy on test data
logistic_accuracy <- mean(logistic_preds_binary == test_data$isB_numeric)
cat("1 (b) Logistic regression accuracy:", logistic_accuracy, "\n")


```

1.(c)
```{r}
logistic_roc <- roc(test_data$isB, logistic_preds)
logistic_auc <- auc(logistic_roc)
cat("1 (c) Logistic regression AUC:", logistic_auc, "\n")

```

1.(d)
```{r}
train_data$isB_factor <- factor(train_data$isB, levels = c("No", "Yes"))
test_data$isB_factor <- factor(test_data$isB, levels = c("No", "Yes"))
#cross-validation manually for different cp values
set.seed(2024)
cv_results <- lapply(seq(0.01, 0.1, by = 0.01), function(cp_value) {
  model <- rpart(isB_factor ~ . -letter -isB -isB_numeric, data = train_data,
                 method = "class", control = rpart.control(cp = cp_value))
  preds <- predict(model, newdata = test_data, type = "class")
  accuracy <- mean(preds == test_data$isB_factor)
  return(data.frame(cp = cp_value, accuracy = accuracy))
})

cv_results_df <- do.call(rbind, cv_results)

#best cp value based on the highest accuracy
best_result <- cv_results_df[which.max(cv_results_df$accuracy), ]
best_cp <- best_result$cp

#train the final CART model using the best cp
final_cart_model <- rpart(isB_factor ~ . -letter -isB -isB_numeric, data = train_data,
                          method = "class", control = rpart.control(cp = best_cp))

#Predict on the test set and calculate accuracy
cart_preds <- predict(final_cart_model, newdata = test_data, type = "class")
cart_accuracy <- mean(cart_preds == test_data$isB_factor)

cat("\n1 (d) Selected ccp_alpha:", best_cp, "\n")
cat("1 (d) CART model accuracy on the test set:", cart_accuracy, "\n")

```
I ran different CART models using various ccp values and checked how well each model performed on the test set. I picked the ccp_alpha value that gave the highest accuracy. The final CART model, using ccp_alpha of 0.01, had an accuracy of about 92.62% on the test set.


1.(e)
```{r}
train_data <- train_data[, !names(train_data) %in% c("Unnamed: 0")]
test_data <- test_data[, !names(test_data) %in% c("Unnamed: 0")]
#Random Forest model
set.seed(2024)
rf_model <- randomForest(isB_factor ~ . -letter -isB -isB_numeric, data = train_data)

#Predict on the test set
rf_preds <- predict(rf_model, newdata = test_data)

#accuracy on the test set
rf_accuracy <- mean(rf_preds == test_data$isB_factor)
cat("\n1 (e) Random Forest model accuracy on the test set:", rf_accuracy, "\n")

```

1.(f)
random forest has the highest accuracy. accuracy is more important because it ensures the model makes correct predictions. Interpretability is less important here since the main goal is to get the highest number of correct results.

2.(a)
```{r}
most_frequent_class <- names(sort(table(train_data$letter), decreasing = TRUE))[1]
baseline_accuracy_multi <- mean(test_data$letter == most_frequent_class)
cat("2 (a) Baseline accuracy for multi-class:", baseline_accuracy_multi, "\n")

```


2(b)
```{r}
#Clean constant column
constant_columns <- sapply(train_data[, -which(names(train_data) %in% c("letter", "isB", "isB_numeric", "isB_factor"))], function(col) length(unique(col)) == 1)
constant_columns <- names(train_data)[constant_columns]
cat("Constant columns:\n")
train_data_filtered <- train_data[, !names(train_data) %in% constant_columns]
test_data_filtered <- test_data[, !names(test_data) %in% constant_columns]

#LDA model
lda_model <- lda(letter ~ . -isB -isB_numeric -isB_factor, data = train_data_filtered)
#Predict on test data
lda_preds <- predict(lda_model, newdata = test_data_filtered)$class
lda_accuracy <- mean(lda_preds == test_data$letter)

cat("2 (b) LDA model accuracy on the test set:", lda_accuracy, "\n")

```

2(c)
```{r}
set.seed(2024)
cart_model_multi <- train(letter ~ . -isB, data = train_data, method = "rpart",
                          trControl = trainControl(method = "cv", number = 5))

#Predict on test data
cart_preds_multi <- predict(cart_model_multi, newdata = test_data)
cart_accuracy_multi <- mean(cart_preds_multi == test_data$letter)
cat("2 (c) CART model accuracy for multi-class:", cart_accuracy_multi, "\n")

```
I used 5-fold cross-validation to train multiple CART models and find the best cp value that provided the highest average accuracy. The CART model selected through this process achieved an accuracy 0.7475936 on the test set.


2(d)

```{r}
train_data$letter <- as.factor(train_data$letter)
test_data$letter <- as.factor(test_data$letter)
set.seed(2024)
bagging_model <- randomForest(letter ~ . -isB -isB_numeric -isB_factor, data = train_data)

#Predict on the test set
bagging_preds <- predict(bagging_model, newdata = test_data)
#Calculate accuracy on the test set
bagging_accuracy <- mean(bagging_preds == test_data$letter)
cat("2 (d) Bagging model accuracy:", bagging_accuracy, "\n")


```

2(e)

```{r}
set.seed(2024)
rf_cv_model <- train(letter ~ . -isB -isB_numeric -isB_factor, data = train_data, method = "rf",
                     trControl = trainControl(method = "cv", number = 3),  
                     tuneGrid = expand.grid(mtry = c(2, 4, 6)))  #select less fold and range for faster speed of running time in r

#Predict on the test set
rf_cv_preds <- predict(rf_cv_model, newdata = test_data)
rf_cv_accuracy <- mean(rf_cv_preds == test_data$letter)
cat("2 (e) Random Forest with cross-validation accuracy:", rf_cv_accuracy, "\n")


```



2(f)
```{r}
set.seed(2024)
gbm_model <- gbm(letter ~ . -isB -isB_numeric -isB_factor, data = train_data, distribution = "multinomial",
                 n.trees = 300,  
                 interaction.depth = 15,  
                 cv.folds = 3,  
                 n.cores = 1) 

#Predict on the test set
gbm_preds <- predict(gbm_model, newdata = test_data, n.trees = 300, type = "response")
gbm_final_preds <- colnames(gbm_preds)[apply(gbm_preds, 1, which.max)]

#Calculate accuracy on the test set
gbm_accuracy <- mean(gbm_final_preds == test_data$letter)
cat("2 (f) Gradient Boosting accuracy:", gbm_accuracy, "\n")


```
